\chapter{Methodology}\label{methodology}

The research methodology comprises multi-stage pipeline. It will
systematically process interoperability protocols from the technical
specifications towards the formal mathematical verification. Every stage
in this methodology is treated as sequential and dependent; the success
of every following stage directly follows from the output of the
previous stage. This ensures the logical connection of the whole
process.

\section{Formal Specification of Interoperability
Protocols}\label{formal-specification-of-interoperability-protocols}

The goal of this stage is to create a strict representation of
representative protocols using algebraic language. This formalisation is
important to remove details of their implementation that have no
connection to the semantic behaviour. This allows comparison between
asset-based and data-based mechanism.

\subsection{Select Representative
Protocols}\label{select-representative-protocols}

Protocols that have wide adoptation and formal documentation will be
used. For asset-based, Hashed Time-Lock Contracts (HTLC) is considered
because it serves as the de-facto guideline for atomic swap. For
data-based, Oracle-based Message Relay is considered because of its
fundamental paper in delivering external data towards blockchains.

\subsection{Identify Actions}\label{identify-actions}

Identify the elementary operations that makes up each protocol:
\textbf{Asset-Based \((A_{asset})\):} \emph{lock} (locking of asset),
\emph{unlock} (moving asset out using secret), \emph{refund} (returning
of token after a timeout). \textbf{Data-based \((A_{data})\):}
\emph{request} (asking of data), \emph{verify} (confirmation of data
integrity), \emph{relay} (delivery of proven message to destination
chain). Let: \[ D: \text{set of data (assets or messages)} \]
\[ P: \text{set of ports (e.g., source chain, relayer, destination chain)} \]
Then: \[ A = {s_p(d), r_p(d) | d \in D, p \in P} \] Protocol specific
actions are mapped into this structure: Asset-based:
\[ lock \rightarrow s_p(d) \]
\[ unlock \rightarrow c_p(d) \text{ (succesful transfer)} \]
\[ refund \rightarrow s_p(d) \text{ (reverse flow)} \] Data-based:
\[ request \rightarrow s_p(d) \]
\[ verify \rightarrow \text{ internal communication} \]
\[ relay \rightarrow c_p(d) \]

\subsection{Encode Process}\label{encode-process}

Form the logical protocol using the following operators:
\textbf{Sequential (\(\cdot\)):} To signify strict sequential process.
\textbf{Alternative (\(+\)):} To model branching of decision.
\textbf{Parallel (\(||\)):} To model interaction between two different
blockchain ledger. System components (e.g., sender, channel, receiver)
are modelled as separate processes and composed as:
\[ S || K || R || L \]

\subsection{Encapsulation and
Abstraction}\label{encapsulation-and-abstraction}

To control system visibility: \textbf{Encapsulation set:}
\[ H = {s_p(d), r_p(d) | p \text{ is internal}} \] Ensures that
unmatched actions are blocked. \textbf{Abstraction set:}
\[ I = {c_p(d) | p \text{ is internal}} \] Internal communications are
hidden as silent steps (\(\tau\)).

\subsection{Final System
Representation}\label{final-system-representation}

The complete system can be expressed as:
\[ \tau_I(\partial_H(S || K || R || L)) \] This serves as the
implementation model (IMP) that will be used for analysis.

\section{Semantic Extraction Using Labelled Transition
System}\label{semantic-extraction-using-labelled-transition-system}

The goal of this stage is to give the process terms made from the
preceding stage formal interpretation using Labelled Transition System
(LTS). Through this, every algebraic term will have its corresponding
dynamic model that shows all possible observable behaviour of the
system.

\subsection{Lay out Structural Operational Semantics
(SOS)}\label{lay-out-structural-operational-semantics-sos}

SOS will be used in order to see the meaning of every operator in the
process algebra using inference rules. Every rule is composed of premise
and one conclusion: \[ \frac{premises}{conclusion} \]

\subsection{Construct the LTS Model}\label{construct-the-lts-model}

For every specific protocol, LTS will be formed as a triple
\[ (S, A, \rightarrow) \]:

\textbf{States \((S)\):} set of states comprising the original process
term and all terms that can be achieved from it (reachable states).
\textbf{Labels \((A)\):} labels of transition that refers to atomic
actions that the system executes. \textbf{Transition \((\rightarrow)\):}
relation of transition \((s \xrightarrow{a} s')\) that shows change from
one state to the another after the action \((a)\). \textbf{pLTS} will be
used when probabilisq behavious is included: captures message loss,
corruption, and nondeterminism.

\subsection{Extract Traces at Reachable
States}\label{extract-traces-at-reachable-states}

\textbf{Reflexive States:} reflexive transitive closure
\((\rightarrow*)\) to identify all possible states that can be reached
by the protocol from the starting state. \textbf{Traces:} get the finite
sequences of actions \[ a_1, a_2, ..., a_n \] that the system executed
in order to illustrate the totally of its oberservable behaviour.

\section{Description of Observable
Behaviour}\label{description-of-observable-behaviour}

The goal of this stage is to formally identify the oberservable
behaviour of the interoperability protocols using the extraction of
traces and execution paths from the LTS mdoels. This survey will serve
as the basis of comparison of asset-based and data-based mechanims at
the noticeable level of their connecition.

\subsection{Give meaning to
Observable}\label{give-meaning-to-observable}

The behaviour of a system is tread as oberservable if this comprises
actions that can be seen or experience by an outside observer or
adjacent system by means of communication. An observable behaviour is
formally defined as an external specification (SPEC) that remains after
applying the \textbf{abstraction operator (\(\tau_I\))} in an
implementation (IMP) TWOSIMPLEPROTCOLS. In this methodology, the
internal connections between chains are treated as \textbf{hidden
actions} and are made as silent steps (\(\tau\)) in order to only see
the connection of input and output between the system and its
environenment . The rigour of the observation does not come from the
absence of \(\tau\) but on how abstraction lessens the complexity of the
internal structure towards a simple SPEC that affirms the correctness of
interoperability.

Internal calculations or local changes in the storage that cannot be
seen at cross-chain communication will not be included in this survery.

\subsection{Extract Traces and Execution
Paths}\label{extract-traces-and-execution-paths}

\textbf{Traces:} Traces (\(Tr(q)\)) will be extracted by listing all
possible series of oberservable actions that can be execute by the
protocol from its origin state (\(q_0\)). A trace is a string
\(\sigma \in A^{}*\) where there persist a sequential transitions
towardsa new state. \textbf{Execution Path:} In order to survey the
structure of the protocol deeper, execution paths or the total flow of
the transition \((s \xrightarrow{a} s')\) that shows not only the series
of actions but also points where there is nondeterministic choice or
branch of decision.

\textbf{Justify as Basis of Interoperability:} The oberservable
behaviour is the fundamental basis of interoperability because the
communication between two blockchain is being treated as a black box
interaction. In a chain\textquotesingle s perspective, the internal
implementation of the other chain has no corelation; what only matters
is the actions being sent and being accepted by it in order to make sure
the semantic consistency. In other words, the correctness of
interoperability depends on whether the observable traces of the two
system matches under the one specified protocol.

\section{Evaluation Criteria}\label{evaluation-criteria}

The goal of this stage is to set the formal guideline that will be used
in comparing asset-bsed and data-based protocols. By means of
identigying the level of similarity and difference of their behaviour,
results of the analysation under one concept of semantic
interoperability can be given interpretations.

\subsection{Similarity, Difference, and
Equivalence}\label{similarity-difference-and-equivalence}

\textbf{Equivalence:} Two protocol is equivalent if their process graph
shows equal pattern of interaction to their environment. Formally, this
is identified as an equivalence class (\(|x| = {y : x \sim y}\)) wherein
the system belonging to this has one oberservable behaviour.
\textbf{Similarity:} This refers to having some parts of the trace
matching even if the structure of the process is not fully similar.
\textbf{Difference:} This surfaces when a protocol has branching
structure or deadlock behaviour that cannot be found in another even if
they have matching strings of action.

\subsection{Evaluation Criterias}\label{evaluation-criterias}

\textbf{Trace Equivalence ( \(=_{tr}\) ):} This criteria will be used to
survey if the protocols can execute exactly the same series or strings
of atomic action. Nonetheless, trace equivalence is considered as coarse
because this does not capture the difference from deadlock or the timing
of choices. \textbf{Bisimulation Equivalence ( \(\sim\) ):} This the
stricter criteria (finest). Under this criteria, two protocol need to
have the same branching structure; meaning, in every step of a protocol,
the other should respond to the same action and reach a state that is
again equal.

\subsection{Interpretation of the
Results}\label{interpretation-of-the-results}

The presence of bisimulation between asset-based and data-based protocol
means that the two has semantic interoperability. That is, the meaning
and context of their operation still remains consistent in every
transition.

If the protocols are trace equivalent only but not bisimilar, the
interpretation will have different level of liveness and safety; for
example, one fo them can lead to deadlock at situations where the other
one have not experienced yet.

The adherence to this formal criterias ensures that interoperability
does not only depend on technical connection but also to mathematical
correctness and semantic integrity of the system.
